# Reproducible Research Course Project 1

## Load Necessary Packages
```{r}
library(dplyr)
library(ggplot2)
library(knitr)
library(markdown)
library(lattice)
```

## Loading the Data
```{r}
data <- read.csv("activity.csv")
```

## Calculate Mean Number of Steps per Day
```{r}
steps <- tapply(data$steps, data$date, FUN=sum, na.rm=TRUE)
qplot(steps, binwidth=500, xlab = "Steps per Day")
mean(steps, na.rm=TRUE)
median(steps, na.rm=TRUE)
```

## Find Average Daily Activity Pattern
```{r}
average <- aggregate(x=list(steps=data$steps), by=list(interval=data$interval), FUN=mean, na.rm=TRUE)
ggplot(data=average, aes(x=interval, y=steps)) +
    geom_line() +
    xlab("Time of Day") +
    ylab("Steps Taken")
```

## Impute Missing Values
```{r}
repl <- function(steps, interval) 
{
  replaced <- NA
  if (!is.na(steps))
    replaced <- c(steps)
  else
      replaced <- (average[average$interval==interval, "steps"])
  return(replaced)
}
newdata <- data
newdata$steps <- mapply(repl, newdata$steps, newdata$interval)
stepss <- tapply(newdata$steps, newdata$date, FUN=sum)
qplot(stepss, binwidth=500, xlab="Steps per Day")
mean(stepss)
median(stepss)
```

The mean and median are both higher after removing the NA values from the dataset.
These values were being replaced with zeros, which artificially reduced the calculated mean and median values. 
While we still do not know what these values, should be, we replaced these zeroes in the calculations, with the averages of the non-NA values, which is what caused these calculations to increase the second time around.

```{r}
partofweek <- function(d) 
{
    days <- weekday(d)
    ifelse (days == "Saturday" | days == "Sunday", "weekend", "weekday")
}

dow <- sapply(newdata$date, partofweek)
newdata$pow <- as.factor(dow)

dataa <- aggregate(steps ~ pow+interval, data=newdata, FUN=mean)

xyplot(steps ~ interval | factor(pow), layout = c(1, 2), xlab="Interval", ylab="Steps", type="l", lty=1, data = newdata)
```